Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network

### 2017-10-27 www-2018/source/paper.tex
Intro edits.

\begin{abstract}
It is desirable for critical large-scale communication networks
such as the internet to be resilient against any types of disruption
\cite{sterbenz_resilience_2010}.
Vulnerabilities in communication infrastructure can be used by
state, corporate, and private actors to achieve censorship and surveillance,
to the detriment of free expression and free association
\cite{}.
Self-organized networks such as the internet's router network typically have
heavy-tailed degree distributions
\cite{barabasi_scale-free_2009},
making them resilient against random faults,
but highly vulnerable to targeted attacks against high-centrality nodes
\cite{albert_error_2000}.
Cryptographic solutions fail to address the underlying topological problem,
and remain vulnerable to man-in-the-middle attacks
\cite{nayak_different_2010}
and coercion
\cite{grewal_caveat_2013}.
Coercion-resistant, topological approaches to attack tolerance are needed
to address the current vulnerability of communications infrastructure to
targeted censorship and surveillance.
We present a novel concurrent multipath routing (CMR) algorithm for the wraparound
butterfly network topology,
as well as a highly attack-tolerant
Structured Multipath Fault Tolerance (SMFT) architecture
which incorporates the butterfly CMR algorithm.
Our architecture is the first theoretical demonstration of a
coercion-resistant point-to-point communication architecture
that does not require infinitely transitive trust.
We also present for modeling attack-tolerance as a function of
trust transitivity,
and show how the internet's router network could be rewired for greater
attack-tolerance.
We identify a previously unexplored relationship between
network topology, trust transitivity, and attack-tolerance,
and provide a framework for further exploration of this relationship.
Our results are the first we're aware of to demonstrate
that large-scale communication networks can be
designed to tolerate targeted attacks,
including coercion and other non-technical attacks.
Furthermore, if attack-tolerance requirements are strict enough to justify
a high level of control over network topology, our results could be applied
directly.
More broadly, we hope that this work will serve as a starting point for the
development of additional topology-based attack-tolerant
communication architectures to guard
against the dangers of censorship and surveillance.
\end{abstract}

@@

Is it possible for any large-scale communication network to resist targeted attacks?
The internet was originally designed to withstand targeted (nuclear) attacks \cite{baran_distributed_1964},
and the resilience of the internet has long been part of common wisdom \cite{elmer-dewitt_first_1993}.
But 17 years after Albert et. al \cite{albert_error_2000}
showed the internet's router network to be vulnerable to targeted attacks (vs. random faults),
the fundamental problem remains unsolved.
The ongoing vulnerability of the internet is evidenced by a long history of 
censorship and surveillance incidents achieved by means of targeted attacks
\cite{dainotti_analysis_2011}.
In this paper, we present a theoretical infrastructure for networked communication,
that tolerates not only random faults, but also targeted attacks.
While not yet practical, our proposal is the first demonstration that
attack-tolerant communication networks are even possible.

Much of the world's infrastructure is networked: power grids,
cellular networks, roads, and of course, the Internet.
As with any critical infrastructure, the cost of failures can be
immense, so methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
both theoretical results and recent events (described below)
have demonstrated that the Internet can be surprisingly vulnerable to attack.
Decentralization remains a promising approach towards
building resilient networks,
but there is a need to better understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a router misconfiguration allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without even trying only highlights
the Internet's vulnerability to adversarial faults.

@@

Analysis of the Internet's router network has shown that while it
is remarkably resilient against random faults,
it is highly susceptible to adversarial faults \cite{albert_error_2000}.
These results have been attributed to the scale-free structure of the Internet's
router network
\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
In scale-free networks and other networks with heavy-tail degree distributions,
random failures are highly likely to
affect only low-degree nodes, thus having
little effect.
However, Adversarial faults target the few high-degree nodes,
and therefore remove a large number of edges with each fault.
So while the {\em protocols} of the Internet are decentralized,
the {\em network structure} is somewhat centralized. 
In other words, the protocols of the Internet do not {\em require}
centralization, but centralization may still emerge from the sociotechnical
processes that create its network structure.

With strong theoretical and historical evidence that centralized
network structure can create vulnerabilities,
methods for analyzing structural vulnerabilities and for designing
fault-tolerant networks are needed.
This paper presents several contributions towards advancing those goals.
While our motivation comes from the Internet router network and
World Wide Web,
our work is theoretical, focusing on abstract networks,
and could potentially be applied to many different types of networks,
whether made of physical wires, virtual tunnels,
or other types of links.

We consider a setting in which a source node (Alice)
in a network attempts to route a message
to a target node (Bob) by forwarding it through the links of the network.
A ``link'' in this context could represent any kind of connection
(e.g., physical cables, encrypted channels).
We assume that some nodes in the network may be compromised by an attacker
(Mal).
We also assume that Mal is an adversary of Alice specifically and targets
nodes strategically with the goal of interfering with Alice's communications
(rather than disrupting the network as a whole).
This assumption applies to scalable network architectures that can be made
large enough that an attacker must focus their resources in order to have
a significant effect.
Compromised nodes may behave incorrectly by blocking, altering,
or incorrectly routing messages.
We assume that Mal has full knowledge of the network structure, but has
limited resources and thus can only compromise a fixed number of nodes.

We also assume that nodes {\em trust} their immediate neighbors.
So Mal is unable to compromise Alice's node, or her direct neighbors.
In the commonly used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003},
we would extend that trust transitively to the entire network.
However, we make a weaker and more realistic assumption:
that trust is extended transitively to nodes within a fixed
number of hops.
So only the beginning and end of a path between Alice and Bob is trusted
against interference from Mal.
We call this assumption
{\em bounded trust transitivity}.

@@

It is important to note that the butterfly is a highly structured and constrained
network topology,
very different form those found in social networks and other
self-organized networks.
The reader may wonder whether it is realistic or useful to assume such control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential for an attacker to exploit it through coercion,
without needing to compromise the technology.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.
Additionally, there may be ways to achieve improved attack-tolerance from architectures
more flexible than the butterfly, which is a potential area for future work.

>>>>

\begin{abstract}
It is crucial for large-scale communication networks such as the internet
to be resilient against attacks\cite{sterbenz_resilience_2010}
such as censorship and surveillance,
which pose a threat to free expression and free association.
Self-organized networks such as the internet’s router network typically have
heavy-tailed degree distributions\cite{barabasi_scale-free_2009},
making them highly vulnerable to targeted attacks against central nodes
\cite{albert_error_2000}.
While cryptographic solutions exist,
they fail to address the underlying topological problem,
and remain vulnerable to
man-in-the-middle attacks \cite{nayak_different_2010}
and coercion \cite{grewal_caveat_2013}.
Coercion-resistant, topological approaches to attack tolerance are needed to
address the current vulnerability of communications infrastructure to
censorship and surveillance.
We present a novel concurrent multipath routing (CMR) algorithm for the
wraparound butterfly network topology,
as well as a highly attack-tolerant Structured Multipath Fault Tolerance (SMFT)
architecture which incorporates the butterfly CMR algorithm. We also identify a
previously unexplored relationship between network topology, trust transitivity,
and attack-tolerance, and provide a framework for further exploration of this
relationship. Our work is the first theoretical demonstration of a point-to-point
communication network architecture that can resist coercion and other
non-technical attacks, without requiring infinitely transitive trust.
To address cases where the network structure cannot be fully controlled,
we demonstrate how a snapshot of the internet’s router network can be partially
rewired for greater attack-tolerance.
More broadly, we hope that this work will serve as a starting point for the
development of additional topology-based attack-tolerant communication
architectures to guard against the dangers of censorship and surveillance.
\end{abstract}

@@

Is it possible for any large-scale communication network to resist targeted
attacks?
The internet was originally designed to withstand targeted (nuclear) attacks
\cite{baran_distributed_1964},
and the resilience of the internet has long been part of common wisdom
\cite{elmer-dewitt_first_1993}.
But 17 years after Albert et. al \cite{albert_error_2000}
showed the internet's router network to be vulnerable to targeted attacks
(vs. random faults), the fundamental problem remains unsolved.
The ongoing vulnerability of the internet is evidenced by a long history of 
censorship and surveillance incidents achieved by means of targeted attacks
\cite{dainotti_analysis_2011}.
In this paper, we present the first theoretical architecture
for point-to-point networked communication,
that tolerates not only random faults, but also targeted attacks,
without relying on infinitely transitive trust
\cite{christianson_why_1997}.

Methods for tolerating various kind of faults within networks are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
{\em Adversarial faults},
those in which an adversary can target attacks strategically,
deserve special attention.
Such attacks are both extremely difficult to guard against and 
often have important social implications.
In particular, censorship and surveillance are often achieved
by targeting central network locations and either blocking or capturing
the information flowing through them.
The Internet's decentralized design was motivated
by the need to withstand targeted attacks, such as nuclear strikes
\cite{baran_distributed_1964}.
But despite longstanding common wisdom \cite{elmer-dewitt_first_1993},
both theoretical results and recent events
have demonstrated that the internet is surprisingly vulnerable to attack.

Analysis of the internet's router network has shown that while it
is remarkably resilient against random faults,
it is highly susceptible to adversarial faults \cite{albert_error_2000}.
These results have been attributed to the heavy-tailed
degree distribution of the Internet's router network
\cite{barabasi_emergence_1999,barabasi_scale-free_2009}.
Random failures are highly likely to affect only low-degree nodes,
thus having little effect.
However, adversarial faults target the few high-degree nodes,
and therefore remove a large number of edges with each fault.
So while the {\em protocols} of the Internet are decentralized,
the {\em network structure} is somewhat centralized. 
In other words, the protocols of the Internet do not {\em require}
centralization, but centralization may still emerge from the sociotechnical
processes that create its network structure.

The internet's vulnerability to censorship and other targeted attacks
has been demonstrated by several recent events.
In 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
The action was initiated by government order in Pakistan,
and spread beyond Pakistan when a router misconfiguration allowed the false
routing information to propagate globally.
While the government order and router misconfiguration initiated the outage,
it was the structure of the Internet's router network that allowed a fault in a
single router to propagate globally.
And while the action was not an intentional attack against the global Internet,
the ability of an attacker to succeed without even trying only highlights
the internet's vulnerability to adversarial faults.

@@

Coercion-resistant, topological approaches to attack tolerance are needed to
address the current vulnerability of communications infrastructure
to censorship and surveillance.
This paper presents several contributions towards advancing those goals.
While our motivation by the network of internet routers,
our work is theoretical,
and could potentially be applied to many different types of networks,
whether made of physical wires, virtual tunnels,
or other types of links.

We consider a setting in which a source node attempts to route a message
to a target node, while an adversary attempts to block or intercept the message
by compromising a number of intermediate nodes.
We also assume that the source and destination nodes trust their neighbors
i.e., that the adversary is unable to compromise them,
as in the web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}.
Unlike the web of trust approach, which assumes that trust is infinitely
transitive,
we assume {\em bounded trust transitivity}.

@@

It is important to note that the butterfly is a highly structured and constrained
network topology,
very different form those found in social networks and other
self-organized networks.
The reader may wonder whether it is realistic or useful to assume such control over
the network structure.
We have already seen that whenever a single point of failure exists in the network,
there is a potential for an attacker to exploit it through coercion.
So, {\em attack-tolerance cannot be achieved without the ability to influence
network structure}.
Luckily, there are scenarios in which network topology can be dictated.
Examples include overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
In general,
{\em when the need for attack-tolerance is high enough to warrant investment
in infrastructure, networks can be engineered and maintained as infrastructure}.
It is also worth noting that attack-tolerant networks may be sub-components of
larger, less-constrained systems.
For example, a single server might be replaced by a distributed network of servers,
each with different ownership, physical location, and legal jurisdiction,
without placing any unrealistic constraints on the clients connecting to
those servers.
Additionally, there may be ways to achieve improved attack-tolerance from architectures
more flexible than the butterfly, which is a potential area for future work.
